---
title: 'Coding challenge 5: data wrangling'
author: "Sarah Batten"
date: "2025-03-20"
output: 
  pdf_document:
  html_document:
    toc: true
    toc_float: true
  word_document:
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coding challenge 5

### Question 1
Reading the data into R: 
```{r}
diversity <- read.csv("DiversityData.csv")
meta <- read.csv("Metadata.csv", na.strings = "na")
```

### Question 2
Joining dataframes together: 
```{r}
library(tidyverse) # loading required package
alpha <- left_join(diversity, meta, by = "Code")
```

### Question 3
Finding Pielou's evenness:
```{r}
alpha_even <- alpha %>%
  mutate(even = shannon/log(richness))
```

### Question 4
Finding the mean and standard error evenness grouped by crop over time:
```{r}
alpha_average <- alpha_even %>%
  group_by(Crop, Time_Point) %>%
  summarise(Mean.even = mean(even),
            n = n(),
            sd.dev = sd(even)) %>%
  mutate(std.error = sd.dev/sqrt(n))
```
### Question 5 
Calculating the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column.

```{r}
alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, Mean.even) %>%
  pivot_wider(names_from = Crop, values_from = Mean.even) %>% # pivot to wide format
  mutate(diff.cotton.even = Soil - Cotton) %>%
  mutate(diff.soybean.even = Soil - Soybean)
```

### Question 6
Making plots:

```{r}
alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff") %>%
  ggplot(aes(x = Time_Point, y = value, color = diff)) + # changing color to 'diff' variable
  geom_line() +
  xlab("Time (hrs)") +
  ylab("Difference from soil in Pielou's evenness")
```

### Question 7
Link to GitHub: 
[Coding Challenge 5 Directory](https://github.com/sarahbatten/PLPA6820/tree/main)